---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

medapps2021 <img src = "data-raw/Sticker/hex_medapps2021.png" align = "right" height = 140/>

# medapps2021

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/medapps2021)](https://CRAN.R-project.org/package=medapps2021)
[![R-CMD-check](https://github.com/statsmed-sheep/medapps2021/workflows/R-CMD-check/badge.svg)](https://github.com/statsmed-sheep/medapps2021/actions)
<!-- badges: end -->

`medapps2021` contains tidy format data from [the Association of American Medical College's 2021 FACTS: Applicants and Matriculants Data website](https://www.aamc.org/data-reports/students-residents/interactive-data/2021-facts-applicants-and-matriculants-data). Datasets include the mean entrance exam scores and GPAs of medical school applicants and matriculants for the 2021-2022 admissions cycle across several different demographics.

## Installation

You can install the development version of medapps2021 like so:

``` {r, message = FALSE, results = FALSE}
devtools::install_github("statsmed-sheep/medapps2021")
library(medapps2021)
```

## Data

* `astats_all`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 applicants.
* `mstats_all`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 matriculants.
* `astats_major`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 applicants by undergraduate school major.
* `mstats_major`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 matriculants by undergraduate school major.
* `astats_race`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 applicants by race and ethnicity.
* `mstats_race`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 matriculants by race and ethnicity.
* `astats_sex`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 applicants by sex.
* `mstats_sex`: mean total MCAT exam/sub-section scores and combined/science/non-science GPAs for all 2021-2022 matriculants by sex.

## Example

The tidy format of the data allows the use of tidyverse packages like `dplyr` and `ggplot2` for data wrangling and visualization, respectively.

```{r example, message = FALSE}
library(tidyverse)
library(medapps2021)

# Data wrangling

astats_ex <- astats_all %>% 
  filter(category == "Total Applicants")

# Data visualization

ggplot(astats_ex, aes(x = year, y = num_apps)) + 
  geom_col(position = "dodge") +
  labs(title = "Number of Medical School Applicants, 2018-2022", x = "Admissions Cycle Year", y = "Number of Applicants")
```

## Contributor(s)

Clara Li
